---
title: "Week 2 Project"
author: "Piyush Danej"
date: "15/05/2019"
output:
  md_document:
    variant: markdown_github
---

## Reading the data

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(dplyr)
library(ggplot2)
```

Lets read the data and see at the top 5 rows

```{r}
activity <- read.csv("activity.csv")

print(head(activity))

```

Now that we've read the data , lets address the tasks .

## Tasks

### Make a histogram of the total number of steps taken each day.

```{r}

sumsteps <- with(activity , tapply(steps , date , sum))

plot( as.Date(as.character(names(sumsteps))) , sumsteps , type = "h" ,
        xlab = "Date" , ylab= "Total Steps Taken" ,
         main = "Total Steps taken for each day")

```

### Calculate and report the mean and median of the total number of steps taken per day.


```{r}
meanSteps <-  with(activity , tapply(steps , date , mean , na.rm = TRUE)) 

medianSteps <- with(activity , tapply(steps , date , median, na.rm = TRUE))

allSteps <- data.frame(MeanSteps =  meanSteps , MedianSteps = medianSteps)

```

### Mean and Median Steps Each day
```{r echo = FALSE ,  result = "asis"}
print(allSteps)
```


### Make a time series plot  of the 5-minute interval and the average number of steps taken, averaged across all days.

```{r}
averageSteps <- with(activity , tapply(steps , interval , mean , na.rm = TRUE))

plot(names(averageSteps) , averageSteps , type = "l")
```


### Interval with Max Average Steps

```{r}
maxSteps <- max(averageSteps)

maxInterval <- names(averageSteps[match(maxSteps , averageSteps)])

```

#### The interval with maximum steps throughout all days is `r maxInterval` .

### Calculate the total number of NAs in the data set

```{r}
nas <- is.na(activity$steps)
nacount <- nrow(activity[nas , ])
```

#### Total number of NA in the dataset is `r nacount`

### Impute the missing values in dataset (NA)

```{r}

alldates <- unique(as.character(activity$date))

activity2 <- data.frame(steps = c(), date = c(), interval= c())

for(d in alldates){
    m <- activity[activity$date == d , 1] %>% mean(na.rm = T)
    logicV <- is.na(activity[activity$date == d , 1])
    if(is.na(m) || is.nan(m))
        activity[activity$date ==d , 1][logicV] <- 0
    else
        activity[activity$date ==d , 1][logicV] <- m
    
    activity2 <- rbind(activity2 , activity[activity$date ==d , ])
    }
navals <- activity2[is.na(activity2)]

```
#### Number of NA values in activity 2 : `r navals`

### Make a plot of total steps taken in a day

```{r}
sumsteps2 <- with(activity2 , tapply(steps , date , sum))

plot( names(sumsteps2) %>% as.character() %>% as.Date() ,
        sumsteps , 
        type = "h" ,
        ylab = "Dates" , 
        xlab = "Total Steps" ,
        main = "Total Steps taken for each day" )


```

### Mean and median of toal steps in new dataframe

```{r}
meanSteps2 <-  with(activity2 , tapply(steps , date , mean  )) 

medianSteps2 <- with(activity2, tapply(steps , date , median))

allSteps2 <- data.frame(MeanSteps =  meanSteps2 , MedianSteps = medianSteps2)

```

### Difference in Old mean and new Mean

``` {r}
plot( names(sumsteps) %>% as.character() %>% as.Date(), meanSteps2 , type="p" , pch = 1 , col = "red")
points(names(sumsteps) %>% as.character() %>% as.Date() , meanSteps , type ="p" , pch = 20, col = "blue")
legend( "topright",legend = c("Original Mean Steps" , "Imputed Mean steps") , col = c("red" , "blue") , pch = c(1 , 20) , )

```

#### As we can see in the plot , the old NA Values are now 0 which are depicted by red circles without fill.

### Weekday Weekend
```{r}
days <- weekdays(as.Date(as.character(activity2$date)))
newdays <-lapply(days , function(x){
            if(x %in% c("Monday" , "Tuesday" , "Wenesday" , "Thursday"))
                "Weekday"
            else
                "Weekend"
        }) %>% unlist()

activity2$days <- as.factor(newdays)

head(activity2)
```

### Panel plot based on Weekday / Weekend

```{r}
avg <- with(subset(activity2 , activity2$days == "Weekday") , tapply(steps , interval, mean))
avg2 <- with(subset(activity2 , activity2$days == "Weekend") , tapply(steps , interval, mean))

plot(names(avg) , avg , type="l" , lwd = 1 , col = "blue" , main = "Weekday Averages")
plot(names(avg2) , avg2 , type = "l"  , lwd = 1 , col = "red" , main = "Weekend Averages")

```
